---
title: "Using specchioR"
author: "Edward P. Morris"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The R package `specchioR` aims to provide simple wrappers to allow interaction with the spectral information system [SPECCHIO](http://www.specchio.ch) Java client. This client provides many functions and allows storage and query of a large database of ground-based
spectrometry. Two main options for databases exist; using the global database, kindly hosted by the University of Zurich or hosting your own database (with setup facilitated
using a virtualisation system).

## System requirements
The package has been developed in UNIX (Ubuntu 16.04 LTS), specifics for other operating systems will be updated as they are tested.

### UNIX
+ Java and Java Development Kit (JDK);
    + check if JRE or JDK is needed!
    + Ubuntu 16.04: `sudo apt-get update; sudo apt-get install default-jdk`
+ SPECCIO Java client;
    + download installer from http://specchio.ch/app_download.php
    + extract installer
    + make `specchio-client-installer.jar` executable
    + double click to install and follow instructions
    + keep note of the instalation directory
+ In R the rJava package is needed (requires JDK?)

## Setup rJava for interaction with the SPECCHIO client
Once all of the requirements are met specchioR needs to know the path to the SPECCHIO directory, if you followed the default option this should be `$HOME/SPECCHIO`. We can permanently store this (in `extdata/specchioR_config.yaml`) as:

```{r setpath}
require(specchioR)
setSpecchioPath("~/SPECCHIO")
getSpecchioPath()
```
We also need to define the server parameters for the SPECCHIO client to connect, here is an example using the (test) global database hosted by Univ. Zurich, the function apends
an entry to the db_config.txt file in the SPECCHIO directory, registry of your `user_name` and `password` is probably easiset done for the first time directly using the Java client.

We can open the client from R using `launchGUI()`. Navigating to 'Database/Create a new user account' opens a dialog that can be used make your user account. Once this is done we can permantly store the connection details as:

```{r setserver}
addServerToConfig(protocol="https",
                  server="v473.vanager.de",
                  port="443",
                  path="/specchio_service",
                  user_name="your_user_name",
                  password="your_password",
                  data_source="jdbc/specchio_test",
                  speccio.path=getSpecchioPath()
                  )
```

## Starting the SPECCHIO client
Once these setup options are stored, we can now call the client ready for making queries:

```{r createclient}
# Connect to the client using the first server definition
sp.client <- createClient(server.id = 1, speccio.path=getSpecchioPath())
```



